Thread Pool in GeckoDB/BOLSTER - Integrating a Thread Pool into a high performance Database System \\

### 2018-06-20 Presentations/Paper/paper_thread_pool.tex
Add Testing and memory check to paper

\subsection{Managing Thread States}
As mentioned before, accessing the same memory location from different thread may lead to race conditions, which means the program behaviour depends on the accessing order of the threads. For example, while one thread updates data at a specific memory location another threads writes into the data, which results to inconsistency. In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation. In this subsection, we focus on atomic operations to manage thread states. We consider the following steps for a thread to update its state. At first, the current state has to be checked and, depending on the current state, the state has to be updated. If these steps are implemented using an if statement and an update function, the complete update procedure is not atomic. Consequently, another thread can update the state between the check of the current state and the update. To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic lib. This function combines the check and update in one atomic procedure. We implemented every thread state access with this function to ensure atomic thread state updates. After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.

@@

\subsection{Testing and Memory Control}
As last aspect of our implementation, we show how we test the thread pool and how we ensure to avoid memory leaks. Since geckoDB already uses GTest \cite{sen2010quick} for unit testing, we also wrote our test using this framework.

>>>>

\subsection{Managing Thread States}
As mentioned before, accessing the same memory location from different thread may lead to race conditions, which means the program behaviour depends on the accessing order of the threads. For example, while one thread updates data at a specific memory location another threads writes into the data, which results to inconsistency. In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation. In this subsection, we focus on atomic operations to manage thread states. We consider the following steps for a thread to update its state. At first, the current state has to be checked and, depending on the current state, the state has to be updated. If these steps are implemented using an if statement and an update function, the complete update procedure is not atomic. Consequently, another thread can update the state between the check of the current state and the update. To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic library. This function combines the check and update in one atomic procedure. We implemented every thread state access with this function to ensure atomic thread state updates. After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.

@@

\subsection{Testing and Memory Check}
As last aspect of our implementation, we show how we test the thread pool and how we ensure to avoid memory leaks. Since geckoDB already uses GTest \cite{sen2010quick} for unit testing, we also wrote our test using this framework. GTest provides a simple, powerful library to perform unit tests in C++. We wrote unit tests for the priority queue, the thread pool itself and the performance monitoring. As mentioned before, we use functions from the stdatomic library of c. Using this library in GTest results in compiler errors with GCC, since the c atomic implementation differs from the c++ one of the standard library and the GCC developers will not fix this issue. Consequently, we wrote a workaround not using the c stdatomic library in the GTest compilation unit.

Since c as low level language does not have an own memory management, memory leaks may occur using heap space allocations. To find and remove all memory leaks in our implementation, we use the tool \emph{valgrind}, which offers different functionalities to find race and jump condition and to count unfreed heap blocks. We found memory leaks in our priority queue tests and also in the deallocation of thread pool instances. In the following sections, we show our experimental setup and the results of our evaluation.
